---
title: "Source Tracking Analysis"
author: "Jincheng Wang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
# Loading library here
library(tidyverse)
library(FEAST)
library(extrafont)
library(biomformat)
library(ape)
library(qiime2R)
sessionInfo()
# devtools::install_github("ETaSky/FEAST", ref = "removewritingfiles")
np = switch(Sys.info()[['sysname']], Windows = 1, Darwin = 4)
```

The source tracking analysis will be performed using FEAST (Shenhav et al. 2019), a newly developed algorithm using a similar logic and idea as the original source tracking by Dan Knight. Jincheng made some changes on the branch FEAST_beta  is used for this analysis.

## Import files
### baby and maternal samples
```{r}
# all body sites to use for now
ss <- c("Feces", "Mouth", "Forearm", "Vagina", "Right_Areola", "Nose")

# baby qza address
bb_qza_dirs <- paste0("../data/split-data/Baby-", ss[1:3])
# mom qza address
mm_qza_dirs <- paste0("../data/split-data/Mom-", ss)

# import qza tables
bb_qza <- parallel::mclapply(bb_qza_dirs, FUN = function(x){read_qza(paste0(x, "/rarefy-table.qza"))}, mc.cores = np)
names(bb_qza) <- paste0("Baby-", ss[1:3])
#lapply(bb_qza, function(x){x$data %>% dim})

mm_qza <- parallel::mclapply(mm_qza_dirs, FUN = function(x){read_qza(paste0(x, "/rarefy-table.qza"))}, mc.cores = np )
names(mm_qza) <- paste0("Mom-", ss)

#lapply(mm_qza, function(x){x$data %>% dim})


# import metadata
mf <- read.delim("../data/processed-data/Metadata_Baby_Seeding_all_samples_final.txt", header = T, sep = "\t", na.strings = c("", "NA"), stringsAsFactors = F)
## extract metadata for source tracking analysis
mf_st <- mf %>% filter(sample_name %in% (sapply(c(bb_qza, mm_qza), FUN = function(x)colnames(x$data)) %>% unlist()))
mf_st$body_site_corrected[mf_st$body_site_corrected=="Right_Forearm"] = "Forearm"
```

### Gauze samples (this needs to be udpated once uploading gauze sequences into qiita)
```{r}
ft_qza_all <- read_qza("~/Dropbox/41-JCHWANG-NYU/Projects/05-BabyProject2017RobKnight/GauzeSeq/3-deblur/SA-P1-P3_deblur_no-mitochondria-no-chloroplast_table.qza")
mf_gauze <- read.table("~/Dropbox/41-JCHWANG-NYU/Projects/05-BabyProject2017RobKnight/GauzeSeq/0-Doc/Mapping_baby_seeding_gauze.txt", header = T, sep = "\t")

ft_tbl_gauze <- ft_qza_all$data[, mf_gauze$sample_name] %>% as.data.frame %>% filter(rowSums(.)>0)
```


## FEAST analysis
```{r}
family_ids <- mf_st$familyid_unique %>% unique()
Feast_outs_ <- list()
for (i in seq(length(family_ids))){
    # i = 5
    ## do a check so that restart the code is possible
    mf_tmp <- mf_st %>% filter(familyid_unique==family_ids[i], (mom_baby=="Baby" | date_sampling_category_days_continuous<=30)) # select baby sample for this family as well as maternal samples that are within 30 days
    
    bb_sinks_no <- length(mf_tmp$body_site_type[mf_tmp$mom_baby=="Baby"])  # no of baby samples
    mm_srcs_no <- length(unique(mf_tmp$body_site_type[mf_tmp$mom_baby=="Mom"]))  # no of available maternal sources
    if(mm_srcs_no==0){
        print(paste0("No maternal source ", family_ids[i]))
        next
    } 
    if(bb_sinks_no==0){
        print(paste0("No baby samples ", family_ids[i]))
        next
    }
    
    # construct feast count table
    ct_tmp <- split(mf_tmp %>% pull(sample_name), list(mf_tmp$mom_baby, mf_tmp$body_site_corrected), sep = "-")
    ct_tmp <- lapply(seq_along(ct_tmp), function(x){c(bb_qza, mm_qza)[[names(ct_tmp)[x]]]$data[, ct_tmp[[x]], drop = F] %>% as.data.frame() %>% filter(rowSums(.)>0) %>% rownames_to_column(var = "ASV")})
    
    # construct feast metadata
    mf_feast_tmp <- mf_tmp %>% select(sample_name)
    mf_feast_tmp$Env <- mf_tmp$body_site_corrected
    mf_feast_tmp$SourceSink <- NA
    mf_feast_tmp$SourceSink[which(mf_tmp$mom_baby=="Mom")] <- "Source"
    mf_feast_tmp$SourceSink[which(mf_tmp$mom_baby=="Baby")] <- "Sink"
    mf_feast_tmp$id <- NA
    mf_feast_tmp$id[which(mf_feast_tmp$SourceSink=="Sink")] <- seq(sum(mf_feast_tmp$SourceSink=="Sink"))
    
    # add gauze samples to CSR babies
    if(mf_tmp$birth_mode_ms[mf_tmp$mom_baby=="Baby"][1]=="CSseed"){
        ## adding gauze sample, randomly select 3 out of 14.
        mf_feast_tmp_gauze <- mf_gauze[sample.int(14, 3), ] %>% select(sample_name)
        mf_feast_tmp_gauze$Env <- "Gauze"
        mf_feast_tmp_gauze$SourceSink <- "Source"
        mf_feast_tmp_gauze$id <- NA
        mf_feast_tmp <- rbind(mf_feast_tmp, mf_feast_tmp_gauze)
        
        ct_tmp$gauze <- ft_tbl_gauze[, mf_feast_tmp_gauze$sample_name, drop = F] %>% filter(rowSums(.)>0) %>% rownames_to_column(var = "ASV")
    }
    
    ct_tmp <- Reduce(function(x, y){full_join(x, y, by = "ASV")}, ct_tmp) %>% column_to_rownames(var = "ASV") %>% t()
    ct_tmp[is.na(ct_tmp)] <- 0
    
    rownames(mf_feast_tmp) <- mf_feast_tmp$sample_name
    
    Feast_outs_[[family_ids[i]]] <- list(ct = ct_tmp, mf = mf_feast_tmp)
}

save(Feast_outs_, file = "tmp_feaast_outs_.Rdata")

Feast_outs <- parallel::mclapply(Feast_outs_, function(x){FEAST(C = x$ct, metadata = x$mf[, -1], EM_iterations = 1000, COVERAGE = 5000, different_sources_flag = 0)}, mc.cores = np)

save(Feast_outs, file = "Feast_out.Rdata")

```

## Summarize
```{r}
#load("Feast_out.Rdata")
length(Feast_outs) # 157 families were included in the final source tracking analysis


tidy_feasts_out <- function(x){
    x_ <- x %>% rownames_to_column(var = "BabySampleSite") %>% gather(key = "Mom_Samples", value = "Prop", -1) %>% separate(Mom_Samples, into = c("Mom_SampleID", "body_site_type"), sep = "_") %>% merge(., mf_st %>% select(sample_name, body_site_corrected), by.x = "Mom_SampleID", by.y = "sample_name", all.x = T) %>% mutate(mom_body_site_corrected = ifelse(Mom_SampleID=="Unknown", "Unknown", ifelse(body_site_type=="Gauze", "Gauze", body_site_corrected)), body_site_corrected = NULL) %>% group_by(BabySampleSite, mom_body_site_corrected) %>% summarise(body_site_ct = n(), body_site_prop = sum(Prop)) %>% ungroup()
    return(x_)
}

Feast_outs_sum <- parallel::mclapply(Feast_outs, tidy_feasts_out, mc.cores = np) # sum the proportions from the same source for each samples

Feast_outs_sum_tbl <- do.call("rbind", Feast_outs_sum) # convert to a dataframe from a list of separated families

Feast_outs_sum_tbl2 <- Feast_outs_sum_tbl %>% filter(mom_body_site_corrected %in% c("Vagina", "Gauze")) %>% mutate(mom_body_site_corrected="Vaginal") %>%  group_by(BabySampleSite, mom_body_site_corrected) %>% summarise(across(body_site_ct:body_site_prop, sum)) %>% rbind(., Feast_outs_sum_tbl %>% filter(!mom_body_site_corrected %in% c("Vagina", "Gauze"))) # this is to merge vagina and gauze as the same source for those CSR babies according to a discussion with Gloria on 10/20/20 

Feast_outs_sum_tbl_ <- Feast_outs_sum_tbl2 %>% separate(BabySampleSite, into = c("BabySampleID"), sep = "_", remove = F) %>% merge(., mf_st, by.x = "BabySampleID", by.y = "sample_name", all.x = T) %>%  # add metadata to the summarized table
    filter(!(birth_mode_ms=="CS" & mom_body_site_corrected=="Gauze")) # remove the gauze attribution of 2 NY CSself families in skin samples
write.csv(Feast_outs_sum_tbl_, file = "../data/Feast_output_complete_tables.csv", quote = F, row.names = F)

Feast_outs_sum_tbl_stats <- Feast_outs_sum_tbl_ %>% filter(date_sampling_category_days_continuous!=2) %>% mutate(Group_day = ifelse(date_sampling_category_days_continuous<=1, 1, date_sampling_category_days_continuous)) %>% group_by(mom_body_site_corrected, body_site_type, Group_day) %>% dunn_test(data = ., body_site_prop ~ birth_mode_ms, p.adjust.method = "none") %>% mutate(p.adj.signif = NULL) # perform dunn_test test at each time point for each body site source and each body site sink

Feast_outs_sum_tbl_stats$p.adj = p.adjust(Feast_outs_sum_tbl_stats$p, method = "fdr") # perform fdr adjustment on all the statistical p-value from the dun_test
Feast_outs_sum_tbl_stats$equal = ifelse(Feast_outs_sum_tbl_stats$p.adj<=0.1, F, T) # determine if null hypothesis (p1=p2) is rejected (F) or not (T)

Feast_outs_sum_tbl_means <- Feast_outs_sum_tbl_ %>% filter(date_sampling_category_days_continuous!=2) %>% mutate(Group_day = ifelse(date_sampling_category_days_continuous<=1, 1, date_sampling_category_days_continuous)) %>% group_by(mom_body_site_corrected, body_site_type, Group_day, birth_mode_ms) %>% summarise(mean = Hmisc::smean.cl.boot(body_site_prop) %>% paste(collapse = ",")) %>% ungroup %>% separate(mean, into = c("mean", "lower", "upper"), sep = ",") %>% mutate(across(mean:upper, ~round(as.numeric(.x), 3))) %>% arrange(mom_body_site_corrected, body_site_type, Group_day, mean)  # calculate the means for each timepoint, this is to assign group letter using the stats from the above analysis

## assigning statistical groups
Tmp <- lapply(list(Feast_outs_sum_tbl_stats, Feast_outs_sum_tbl_means), function(x){split(x, list(x$mom_body_site_corrected, x$body_site_type, x$Group_day))})
for (i in seq(length(Tmp[[1]]))){
    tt = Tmp[[2]][[i]]$birth_mode_ms %>% combn(., 2) %>% apply(2, sort) %>% t() %>% as.data.frame() %>% left_join(., Tmp[[1]][[i]] %>% select(group1, group2, equal), by = c("V1"="group1", "V2"="group2"))
    if(nrow(tt)==3){
        Tmp[[2]][[i]]$p.adj.l = 
            if(tt[1, 3]){
                if(tt[2, 3]){
                    rep(NA, 3)
                }else{
                    if(tt[3, 3]){
                        c("a", "ab", "b")
                    }else{
                        c("a", "a", "b")
                    }
                }
            }else{
                if(tt[3, 3]){
                    c("a", "b", "b")
                }else{
                    c("a", "b", "c")
                }
            }
    }else{
        Tmp[[2]][[i]]$p.adj.l = 
            if(tt[1,3]){c("a", "a")}else{c("a", "b")}
    }
}
Feast_outs_sum_tbl_means <- do.call("rbind", Tmp[[2]])

# 1st month
ggplot(Feast_outs_sum_tbl_ %>% filter(date_sampling_category_days_continuous<=30), aes(x = date_sampling_category_days_continuous, y = body_site_prop, color = mom_body_site_corrected)) +
    stat_summary(geom="ribbon", aes(fill = mom_body_site_corrected), fun.data=mean_cl_boot, fun.args = list(conf.int=0.95), alpha = 0.3, color = NA)+
    stat_summary(geom="line", fun = mean, size = 1)+
    facet_grid(body_site_type~birth_mode_ms, scales = "free_y") +
    theme_bw(base_size = 10) + theme(panel.grid = element_blank(), panel.background = element_rect(color = NA), aspect.ratio = 0.8) +
    labs(x = "Age (day)", y = "Proportions of maternal Source", color = "Maternal Sources", fill = "Maternal Sources") +
    #scale_color_manual(values = ss_clr) +
    #scale_fill_manual(values =  ss_clr) +
    scale_x_continuous(breaks = c(0, 1, 2, 7, 14, 21, 30))
ggsave("../results/FEAST_plot_1mon.pdf", width = 7, useDingbats = F)

Feast_outs_sum_tbl_stats_label <- Feast_outs_sum_tbl_means %>% filter(Group_day<=30) %>% group_by(mom_body_site_corrected) %>% mutate(y.pos = max(upper, na.rm =T)) %>% group_by(mom_body_site_corrected, body_site_type, Group_day) %>% mutate(y.pos = y.pos*c(1.1, 1.2, 1.3))
ggplot(Feast_outs_sum_tbl_ %>% filter(date_sampling_category_days_continuous<=30), aes(x = date_sampling_category_days_continuous, y = body_site_prop, color = birth_mode_ms)) +
    stat_summary(geom="ribbon", aes(fill = birth_mode_ms), fun.data=mean_cl_boot, fun.args = list(conf.int=0.95), alpha = 0.3, color = NA)+
    stat_summary(geom="line", fun = mean, size = 1) +
    geom_text(data = Feast_outs_sum_tbl_stats_label, aes(x = Group_day, y = y.pos, label = p.adj.l), size = 6/.pt) +
    facet_grid(mom_body_site_corrected ~ body_site_type, scales = "free_y") +
    theme_bw(base_size = 10) + theme(panel.grid = element_blank(), panel.background = element_rect(color = NA), aspect.ratio = 0.8) +
    scale_color_manual(values = c("#D0352C", "#67AD57","#497AB4")) +
    scale_fill_manual(values = c("#D0352C", "#67AD57","#497AB4")) +
    scale_y_continuous(n.breaks = 5) +
    labs(x = "Age (day)", y = "Proportions of maternal Source", color = "Birth Mode", fill = "Birth Mode") +
    scale_x_continuous(breaks = c(0, 1, 2, 7, 14, 21, 30))
ggsave("../results/FEAST_plot_1mon_v2.pdf", width = 7, useDingbats = F)

Feast_outs_sum_tbl_2 <- Feast_outs_sum_tbl_ %>% filter(date_sampling_category_days_continuous<=30) %>% group_by(birth_mode_ms, body_site_type, mom_body_site_corrected, date_sampling_category_days_continuous) %>% summarise(Mean = mean(body_site_prop))

ggplot(Feast_outs_sum_tbl_2, aes(x = date_sampling_category_days_continuous, y = Mean, fill = mom_body_site_corrected)) +
    #stat_summary(geom="ribbon", aes(fill = mom_body_site_corrected), fun.data=mean_cl_boot, fun.args = list(conf.int=0.95), alpha = 0.3, color = NA)+
    geom_area()+
    facet_grid(body_site_type~birth_mode_ms, scales = "free_y") +
    theme_bw(base_size = 10) + theme(panel.grid = element_blank(), panel.background = element_rect(color = NA), aspect.ratio = 0.8) +
    labs(x = "Age (day)", y = "Proportions of maternal Source", color = "Maternal Sources", fill = "Maternal Sources") +
    scale_x_continuous(breaks = c(0, 1, 2, 7, 14, 21, 30))

ggplot(Feast_outs_sum_tbl_ %>% filter(date_sampling_category_days_continuous<=30, body_site_type=="Fecal", mom_body_site_corrected=="Vagina"), aes(x = date_sampling_category_days_continuous, y = body_site_prop, color = birth_mode_ms)) +
    stat_summary(geom="ribbon", aes(fill = birth_mode_ms), fun.data=mean_cl_boot, fun.args = list(conf.int=0.95), alpha = 0.3, color = NA)+
    stat_summary(geom="line", fun.y=mean, size = 1)+
    theme_bw(base_size = 10) + theme(panel.grid = element_blank(), panel.background = element_rect(color = NA), aspect.ratio = 0.8) +
    labs(x = "Age (day)", y = "Proportions of maternal Source") +
    scale_x_continuous(breaks = c(0, 1, 2, 7, 14, 21, 30))

# 1st year
ggplot(Feast_outs_sum_tbl_ , aes(x = date_sampling_category_days_continuous, y = body_site_prop, color = mom_body_site_corrected)) +
    stat_summary(geom="ribbon", aes(fill = mom_body_site_corrected), fun.data=mean_cl_boot, fun.args = list(conf.int=0.95), alpha = 0.3, color = NA)+
    stat_summary(geom="line", fun.y=mean, size = 1)+
    facet_grid(body_site_type~birth_mode_ms, scales = "free_y") +
    theme_classic(base_size = 10) + theme(aspect.ratio = 0.8) +
    labs(x = "Age (day)", y = "Proportions of maternal Source", color = "Maternal Sources", fill = "Maternal Sources")
    # scale_color_manual(values = ss_clr) +
    # scale_fill_manual(values =  ss_clr)
ggsave("../results/FEAST_plot_1yr.pdf", width = 7, useDingbats = F)

# day 7 and 30
Feast_outs_sum_tbl_selected <- Feast_outs_sum_tbl_ %>% filter(date_sampling_category_days_continuous %in% c(7, 30))

ggplot(Feast_outs_sum_tbl_selected, aes(x = factor(date_sampling_category_days_continuous), y = body_site_prop, color = birth_mode_ms)) +
    geom_boxplot(outlier.color = NA) +
    geom_point(position = position_jitterdodge(jitter.width = 0.05), alpha = 0.5) +
    facet_grid(body_site_type~mom_body_site_corrected, scales = "free", labeller = labeller(mom_body_site_corrected = function(x){paste0("Maternal: ", x)})) +
    theme_classic(base_size = 10) + theme(aspect.ratio = 1.5) +
    labs(x = "Age (day)", y = "Proportions of maternal Source", color = "Birth Mode")
ggsave("../results/FEAST_plot_day7and30.pdf", width = 7, useDingbats = F)


```



